---
title: "A quick tour of **qcc**"
author: "Luca Scrucca"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: 
  rmarkdown::html_vignette:
    toc: true
    css: "vignette.css"
vignette: >
  %\VignetteIndexEntry{A quick tour of qcc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(fig.align="center",
               fig.width=5, fig.height=4.5,
               dev.args=list(pointsize=8))

knit_hooks$set(par = function(before, options, envir)
  { if(before && options$fig.show != "none") 
       par(mar=c(4.1,4.1,1.1,1.1), mgp=c(3,1,0), tcl=-0.5)
})
```
       
## Introduction

`qcc` is a contributed R package for **statistical quality control charts** which provides:

- Shewhart quality control charts for continuous, attribute and count data
- Cusum and EWMA charts
- Operating characteristic curves
- Process capability analysis
- Pareto chart and cause-and-effect chart
- Multivariate control charts.

This document gives a quick tour of `qcc` (version `r packageVersion("qcc")`) functionalities. It was written in R Markdown, using the [knitr](http://cran.r-project.org/package=knitr) package for production. 

Further details are provided in the following paper:  

> Scrucca, L. (2004) qcc: an R package for quality control charting and statistical process control. *R News* 4/1, 11-17.

For a nice blog post discussing the `qcc` package, in particular how to implement the *Western Eletric Rules* (WER), see http://blog.yhathq.com/posts/quality-control-in-r.html.


```{r, message=FALSE}
library(qcc)
```

## Shewhart charts

### x-bar chart

```{r}
data(pistonrings)
diameter <- with(pistonrings, qcc.groups(diameter, sample))
head(diameter)

q1 <- qcc(diameter[1:25,], type="xbar", newdata=diameter[26:40,])
plot(q1, chart.all=FALSE)
q1 <- qcc(diameter[1:25,], type="xbar", newdata=diameter[26:40,],
          confidence.level=0.99)
```

Add warning limits at 2 std. deviations:
```{r}
q1 <- qcc(diameter[1:25,], type="xbar", newdata=diameter[26:40,], plot=FALSE)
(warn.limits <- limits.xbar(q1$center, q1$std.dev, q1$sizes, 2))
plot(q1, restore.par = FALSE)
abline(h = warn.limits, lty = 3, col = "chocolate")
```

### R chart
```{r}
q2 <- qcc(diameter[1:25,], type="R")
summary(q2)
q3 <- qcc(diameter[1:25,], type="R", newdata=diameter[26:40,])
summary(q3)
```

### S chart
```{r}
q4 <- qcc(diameter[1:25,], type="S")
summary(q4)
q5 <- qcc(diameter[1:25,], type="S", newdata=diameter[26:40,])
summary(q5)
```

### Variable control limits

```{r}
out <- c(9, 10, 30, 35, 45, 64, 65, 74, 75, 85, 99, 100)
diameter2 <- with(pistonrings, qcc.groups(diameter[-out], sample[-out]))
summary(qcc(diameter2[1:25,], type="xbar"))
summary(qcc(diameter2[1:25,], type="R"))
```

### p and np charts

```{r}
data(orangejuice)
attach(orangejuice)
q1 <- qcc(D[trial], sizes=size[trial], type="p")
summary(q1)

summary(qcc(D[trial], sizes=size[trial], type="np"))
```

Remove out-of-control points (see `help(orangejuice)` for the reasons):
```
inc <- setdiff(which(trial), c(15,23))
q2 <- qcc(D[inc], sizes=size[inc], type="p",
          newdata=D[!trial], newsizes=size[!trial]) 
detach(orangejuice)
```

```{r}
data(orangejuice2)
attach(orangejuice2)
names(D) <- sample
q1 <- qcc(D[trial], sizes=size[trial], type="p")
summary(q1)
q2 <- qcc(D[trial], sizes=size[trial], type="p", 
          newdata=D[!trial], newsizes=size[!trial])
summary(q2)
detach(orangejuice2)
```

### c and u charts

```{r}
data(circuit)
attach(circuit)
q1 <- qcc(x[trial], sizes=size[trial], type="c")
summary(q1)
```

Remove out-of-control points (see `help(circuit)` for the reasons)
```{r}
inc <- setdiff(which(trial), c(6,20))
q2 <- qcc(x[inc], sizes=size[inc], type="c", labels=inc, 
          newdata=x[!trial], newsizes=size[!trial], newlabels=which(!trial))
summary(q2)

q3 <- qcc(x[inc], sizes=size[inc], type="u", labels=inc, 
          newdata=x[!trial], newsizes=size[!trial], newlabels=which(!trial))
summary(q3)
detach(circuit)
```

```{r}
data(pcmanufact)
attach(pcmanufact)
q1 <- qcc(x, sizes=size, type="u")
summary(q1)
detach(pcmanufact)
```

###  Continuous one-at-time data 

```{r}
# viscosity data (Montgomery, pag. 242)
x <- c(33.75, 33.05, 34, 33.81, 33.46, 34.02, 33.68, 33.27, 33.49, 33.20,
       33.62, 33.00, 33.54, 33.12, 33.84)
q1 <- qcc(x, type="xbar.one")
summary(q1)
q2 <- qcc(x, type="xbar.one", std.dev = "SD")
summary(q2)
```

### Standardized p chart

In this example we show how to extend the package by defining a new control chart, i.e. a standardized p chart (`type = "p.std"`).

Function to compute group statistics and center:
```{r}
stats.p.std <- function(data, sizes)
{
  data <- as.vector(data)
  sizes <- as.vector(sizes)
  pbar <- sum(data)/sum(sizes)
  z <- (data/sizes - pbar)/sqrt(pbar*(1-pbar)/sizes)
  list(statistics = z, center = 0)
}
```

Function to compute within-group standard deviation:
```{r}
sd.p.std <- function(data, sizes, ...) { return(1) }
```

Function to compute control limits based on normal approximation:
```{r}
limits.p.std <- function(center, std.dev, sizes, conf)
{
  if(conf >= 1) 
    { lcl <- -conf
      ucl <- +conf 
  }
  else
    { if(conf > 0 & conf < 1)
        { nsigmas <- qnorm(1 - (1 - conf)/2)
          lcl <- -nsigmas
          ucl <- +nsigmas }
      else stop("invalid 'conf' argument.") 
  }
  limits <- matrix(c(lcl, ucl), ncol = 2)
  rownames(limits) <- rep("", length = nrow(limits))
  colnames(limits) <- c("LCL", "UCL")
  return(limits)
}
```

Example with simulated data:
```{r}
# set unequal sample sizes
n <- c(rep(50,5), rep(100,5), rep(25, 5))
# generate randomly the number of successes
x <- rbinom(length(n), n, 0.2)
# plot the control chart with variable limits
summary(qcc(x, type="p", size=n))
# plot the standardized control chart
summary(qcc(x, type="p.std", size=n))
```


## Cusum chart

```{r}
data(pistonrings)
attach(pistonrings)
diameter <- with(pistonrings, qcc.groups(diameter, sample))

q1 <- cusum(diameter[1:25,], decision.interval = 4, se.shift = 1)
summary(q1)

q2 <- cusum(diameter[1:25,], newdata=diameter[26:40,])
summary(q2)
plot(q2, chart.all=FALSE)
```

## EWMA


```{r}
data(pistonrings)
attach(pistonrings)
diameter <- with(pistonrings, qcc.groups(diameter, sample))

q1 <- ewma(diameter[1:25,], lambda=0.2, nsigmas=3)
summary(q1)

q2 <-  ewma(diameter[1:25,], lambda=0.2, nsigmas=2.7, 
            newdata=diameter[26:40,]) 
summary(q2)

x <- c(33.75, 33.05, 34, 33.81, 33.46, 34.02, 33.68, 33.27, 
       33.49, 33.20, 33.62, 33.00, 33.54, 33.12, 33.84)
q3 <-  ewma(x, lambda=0.2, nsigmas=2.7)
summary(q3)
```

## Process capability analysis

```{r}
data(pistonrings)
attach(pistonrings)
diameter <- with(pistonrings, qcc.groups(diameter, sample))

q1 <- qcc(diameter[1:25,], type="xbar", nsigmas=3, plot=FALSE)
process.capability(q1, spec.limits=c(73.95,74.05))
process.capability(q1, spec.limits=c(73.95,74.05), target=74.02)
process.capability(q1, spec.limits=c(73.99,74.01))
process.capability(q1, spec.limits = c(73.99, 74.1))
```

## Multivariate Quality Control Charts

Individual observations data:
```{r}
data(boiler)

q1 <- mqcc(boiler, type = "T2.single", confidence.level = 0.999)
summary(q1)

# generate new "in control" data
boilerNew <- MASS::mvrnorm(10, mu = q1$center, Sigma = q1$cov)
q2 <- mqcc(boiler, type = "T2.single", confidence.level = 0.999, 
           newdata = boilerNew, pred.limits = TRUE)
summary(q2)

# generate new "out of control" data
boilerNew <- MASS::mvrnorm(10, mu = 1.01*q1$center, Sigma = q1$cov)
q3 <- mqcc(boiler, type = "T2.single", confidence.level = 0.999, 
           newdata = boilerNew, pred.limits = TRUE)
summary(q3)

# provides "robust" estimates of means and covariance matrix
rob <- MASS::cov.rob(boiler)
q4 <- mqcc(boiler, type = "T2.single", center = rob$center, cov = rob$cov)
summary(q4)
```

## Pareto chart

```{r}
defect <- c(80, 27, 66, 94, 33)
names(defect) <- c("price code", "schedule date", "supplier code", "contact num.", "part num.")
pareto.chart(defect, ylab = "Error frequency")
```


## Cause and effect diagram

```{r}
cause.and.effect(cause=list(Measurements=c("Micrometers", "Microscopes", "Inspectors"),
                            Materials=c("Alloys", "Lubricants", "Suppliers"),
                            Personnel=c("Shifts", "Supervisors", "Training", "Operators"),
                            Environment=c("Condensation", "Moisture"),
                            Methods=c("Brake", "Engager", "Angle"),
                            Machines=c("Speed", "Lathes", "Bits", "Sockets")),
                 effect="Surface Flaws")
```


## Process variation

In the following simulated data are used to describe some models for process variation. For further details see Wetherill, G.B. and Brown, D.W. (1991) *Statistical Process Control*, New York, Chapman and Hall, Chapter 3.

```{r, echo = FALSE}
set.seed(123) # set seed for reproducibility
```


### Simple random variation

$x_{ij} = \mu + \sigma_W \epsilon_{ij}$

```{r}
mu <- 100
sigma_W <- 10
epsilon <- rnorm(500)
x <- matrix(mu + sigma_W*epsilon, ncol=10, byrow=TRUE)
q <- qcc(x, type="xbar")
q <- qcc(x, type="R")
q <- qcc(x, type="S")
```

### Between and within sample extra variation

$x_{ij} = \mu + \sigma_B u_i + \sigma_W \epsilon_{ij}$

```{r}
mu <- 100
sigma_W <- 10
sigma_B <- 5
epsilon <- rnorm(500)
u <- as.vector(sapply(rnorm(50), rep, 10))
x <- mu + sigma_B*u + sigma_W*epsilon
x <- matrix(x, ncol=10, byrow=TRUE)
q <- qcc(x, type="xbar")
q <- qcc(x, type="R")
q <- qcc(x, type="S")
```

### Autocorrelation

$x_{ij} = \mu + W_i + \sigma_W \epsilon_{ij}$  
where $W_i = \rho W_{i-1} + \sigma_B u_i = \sigma_B u_i + \rho \sigma_B u_{i-1} + \rho^2 \sigma_B u_{i-2} + \ldots$,   
and $W_0 = 0$.

```{r}
mu <- 100
rho <- 0.8
sigma_W <- 10
sigma_B <- 5
epsilon <- rnorm(500)
u <- rnorm(500)
W <- rep(0,100)
for(i in 2:length(W))
    W[i] <- rho*W[i-1] + sigma_B*u[i]
x <- mu + sigma_B*u + sigma_W*epsilon
x <- matrix(x, ncol=10, byrow=TRUE)
q <- qcc(x, type="xbar")
q <- qcc(x, type="R")
q <- qcc(x, type="S")
```

### Recurring cycles

Assume we have 3 working turns of 8 hours each for each working day, so $8 \times 3 = 24$ points in time, and at each point we sample 5 units.

$x_{ij} = \mu + W_i + \sigma_W \epsilon_{ij}$  
where $W_i$ ($i=1,\ldots,8$) is the cycle.

```{r}
mu <- 100
sigma_W <- 10
epsilon <- rnorm(120, sd=0.3)
W <- c(-4, 0, 1, 2, 4, 2, 0, -2) # assumed workers cycle
W <- rep(rep(W, rep(5,8)), 3)
x <- mu + W + sigma_W*epsilon
x <- matrix(x, ncol=5, byrow=TRUE)
q <- qcc(x, type="xbar")
q <- qcc(x, type="R")
q <- qcc(x, type="S")
```

### Trends

$x_{ij} = \mu + W_i + \sigma_W \epsilon_{ij}$  
where $W_i = 0.2 i$

```{r}
mu <- 100
sigma_W <- 10
epsilon <- rnorm(500)
W <- rep(0.2*1:100, rep(5,100))
x <- mu + W + sigma_W*epsilon
x <- matrix(x, ncol=10, byrow=TRUE)
q <- qcc(x, type="xbar")
q <- qcc(x, type="R")
q <- qcc(x, type="S")
```

### Mixture

$x_{ij} = \mu_1 p + \mu_2 (1-p) + \sigma_W \epsilon_{ij}$  
where $p = \Pr(\text{Process #1})$.

```{r}
mu1 <- 90
mu2 <- 110
sigma_W <- 10
epsilon <- rnorm(500)
p <- rbinom(50, 1, 0.5)
mu <- mu1*p + mu2*(1-p)
x <- rep(mu, rep(10, length(mu))) + sigma_W*epsilon
x <- matrix(x, ncol=10, byrow=TRUE)
q <- qcc(x, type="xbar")
q <- qcc(x, type="R")
q <- qcc(x, type="S")
```


### Sudden jumps

$x_{ij} = \mu_i + \sigma_W \epsilon_{ij}$  
where $\mu_i$ is the mean of the process for state $i$ ($i=1,\ldots,k)$.

```{r}
mu <- rep(c(95,110,100,90), c(20,35,25,20))
sigma_W <- 10
epsilon <- rnorm(500)
x <- rep(mu, rep(5, length(mu))) + sigma_W*epsilon
x <- matrix(x, ncol=10, byrow=TRUE)
q <- qcc(x, type="xbar")
q <- qcc(x, type="R")
q <- qcc(x, type="S")
```

